Computational methods in landscape ecology

### 2024-07-11 Manuscript.qmd
Update MS and software

---
title: "Computational methods in landscape ecology"
keywords: keyword1, keyword2
author:
  - name: Maximilian H.K. Hesselbarth
    orcid: 0000-0003-1125-9918
    email: mhk.hesselbarth@gmail.com
    affiliations:
      - name: International Institute for Applied Systems Analysis, Biodiversity, Ecology, and Conservation Group, Laxenburg, Austria
    attributes:
      - corresponding: boolean
  - name: Jakub Nowosad
    orcid: 0000-0002-1057-3721
    email: nowosad.jakub@gmail.com
    affiliations:
      - name: Adam Mickiewicz University, Poznan, Poland
  - name: Marti Bosch
    orcid: 0000-0001-8735-9144
    email: marti.bosch@epfl.ch
    affiliations:
      - name: École polytechnique fédérale de Lausanne, Lausanne, Switzerland
  - name: Martin Jung
    orcid: 0000-0002-7569-1390
    email: jung@iiasa.ac.at
    affiliations:
      - name: International Institute for Applied Systems Analysis, Biodiversity, Ecology, and Conservation Group, Laxenburg, Austria
  - name: Rafael Schouten
    orcid: 0000-0002-8380-0884
    email: rafael.schouten@sund.ku.dk
    affiliations:
      - name: Globe Institute, University of Copenhagen, Copenhagen, Denmark
format:
  pdf:
    keep-tex: true
    template-partials: 
      - title.tex
    include-in-header: 
      text: | 
        \usepackage{lineno}\linenumbers
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
papersize: a4paper
mainfont: "Garamond"
fontsize: 11pt
geometry: margin=25mm
indent: 2m
linestretch: 1.5
bibliography: CLER.bib
csl: springer-basic-author-date.csl
---

@@

{{< include Sections/01-introduction.qmd >}}
{{< include Sections/02-spatial-patterns.qmd >}}
{{< include Sections/05-simulations.qmd >}}

>>>>

---
title: "Computational methods in landscape ecology"
keywords: keyword1, keyword2
author:
  - name: Maximilian H.K. Hesselbarth
    orcid: 0000-0003-1125-9918
    email: mhk.hesselbarth@gmail.com
    affiliations:
      - name: International Institute for Applied Systems Analysis, Biodiversity, Ecology, and Conservation Group, Laxenburg, Austria
    attributes:
      - corresponding: boolean
  - name: Jakub Nowosad
    orcid: 0000-0002-1057-3721
    email: nowosad.jakub@gmail.com
    affiliations:
      - name: Adam Mickiewicz University, Poznan, Poland
  - name: Marti Bosch
    orcid: 0000-0001-8735-9144
    email: marti.bosch@epfl.ch
    affiliations:
      - name: École polytechnique fédérale de Lausanne, Lausanne, Switzerland
  - name: Martin Jung
    orcid: 0000-0002-7569-1390
    email: jung@iiasa.ac.at
    affiliations:
      - name: International Institute for Applied Systems Analysis, Biodiversity, Ecology, and Conservation Group, Laxenburg, Austria
  - name: Rafael Schouten
    orcid: 0000-0002-8380-0884
    email: rafael.schouten@sund.ku.dk
    affiliations:
      - name: Globe Institute, University of Copenhagen, Copenhagen, Denmark
format:
  pdf:
    keep-tex: true
    template-partials: 
      - title.tex
    include-in-header: 
      text: | 
        \usepackage{lineno}\linenumbers
        \usepackage[noblocks]{authblk}
        \renewcommand*{\Authsep}{, }
        \renewcommand*{\Authand}{, }
        \renewcommand*{\Authands}{, }
        \renewcommand\Affilfont{\small}
papersize: a4paper
mainfont: "Verdana"
fontsize: 10pt
geometry: margin=25mm
indent: 2m
linestretch: 1.5
bibliography: CLER.bib
csl: springer-basic-author-date.csl
---

@@

{{< include Sections/01-introduction.qmd >}}
{{< include Sections/02-spatial-patterns.qmd >}}
{{< include Sections/05-simulations.qmd >}}
{{< include Sections/08-software.qmd >}}

### 2024-07-11 Manuscript.tex
Update MS and software

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  11pt,
  a4paperpaper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[]{Garamond}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\usepackage[margin=25mm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{lineno}\linenumbers
\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

@@

@@

@@

@@

@@

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paperpaper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
    \setmainfont[]{Verdana}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\usepackage[margin=25mm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

@@

\usepackage{lineno}\linenumbers
\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

\section{Software and programming
languages}\label{software-and-programming-languages}

``Open science'' involves full transparency and data access at all
stages of the research processes and the integrative and
multi-disciplinary field of ecology can massively benefit from this
concept (Hampton et al. 2015). One key aspect of achieving open science
is to use open source software, which ensures transparency and
consequently increasing the (public) trust into the analyses (Powers and
Hampton 2019). Related to this, script-based workflows have the
advantage over workflows involving graphical user interfaces that the
former are shareable and reproducible (Hampton et al. 2015; Powers and
Hampton 2019). Thus, we will focus the following section on script-based
open-source software suitable for landscape ecology. Given the
increasing availability of data and complexity of methods, future
progress within the field of landscape ecology will depend on the
software and analysis tools available (Costanza et al. 2019)

The open-source and script-based R programming languages (R Core Team
2024) was first released in 1995 (Smith 2016) with a focus on
statistical computing and data visualization. However, due to its
extensions (also called packages), the language is popular in many
academic fields including ecology (Lai et al. 2019) and it is suitable
for spatial data analysis or landscape ecology (e.g., Bivand 2020;
Hesselbarth et al. 2021).

@@

\bibitem[\citeproctext]{ref-Bivand2020}
Bivand RS (2020) Progress in the r ecosystem for representing and
handling spatial data. Journal of Geographical Systems.
\url{https://doi.org/10.1007/s10109-020-00336-0}

@@

\bibitem[\citeproctext]{ref-Hampton2015}
Hampton SE, Anderson SS, Bagby SC, et al (2015) The tao of open science
for ecology. Ecosphere 6:1--13.
\url{https://doi.org/10.1890/ES14-00402.1}

@@

\bibitem[\citeproctext]{ref-Hesselbarth2021}
Hesselbarth MHK, Nowosad J, Signer J, Graham LJ (2021) Open-source tools
in r for landscape ecology. Current Landscape Ecology Reports 6:97--111.
\url{https://doi.org/10.1007/s40823-021-00067-y}

@@

\bibitem[\citeproctext]{ref-Lai2019}
Lai J, Lortie CJ, Muenchen RA, et al (2019) Evaluating the popularity of
r in ecology. Ecosphere 10: \url{https://doi.org/10.1002/ecs2.2567}

@@

\bibitem[\citeproctext]{ref-RCoreTeam2024}
R Core Team (2024) \href{https://www.R-project.org/}{R: A language and
environment for statistical computing}. R Foundation for Statistical
Computing, Vienna, Austria

@@

\bibitem[\citeproctext]{ref-Smith2016}
Smith D (2016) Over 16 years of r project history
